EvoSwarm: An Event-based Architecture for Multi-population Optimization Algorithms

### 2020-05-13 draft.tex
Cloud and BBOB added
Experiments 2 and 3 added. Refs #6 and  #9.

\usepackage{lineno,hyperref}
\usepackage{rotating}
\usepackage{algpseudocode,algorithm}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{xcolor}

@@

For local experiments we have deployed the EvoSwarm application in a
Desktop PC with an AMD Ryzen 9 3900x 12-core processor with 24 threads 
and 16 GB RAM. We used Docker version 19.03.3, build a872fc2f86, and docker-compose version 1.21.0, 
in Ubuntu Linux 18.04; coce is run using Python 3.7.5. Container images and the docker-compose file are available at
\url{https://hub.docker.com/u/mariosky}, and \url{https://github.com/mariosky/EvoSwarm} respectively. 
The BOBB experiments were performed with COCO \cite{hansen2016coco} version bbob.v15.03 in Python, 
the plots were produced with version 2.3.2. % Are we using BBOB plots? - JJ

@@

@@

% % Experiment 1

% Having a multi-population based algorithm can decrease the execution time
% because of the parallel execution of the evolution. But, having heterogeneous 
% populations might enhance evolutionary search and needless
% evaluations than homogeneous systems; heterogeneous settings, if
% done right, increases the diversity of the whole population
% \cite{araujo2008multikulti}.

% But this is a rule of thumb, and it will depend on the degree of heterogeneity,
% as well as on the algorithm itself. Some level of heterogeneity can be
% implemented by just changing the configuration parameters of each population,
% but in this case, we are interested in a heterogeneous search strategy. 
% Therefore, in this experiment we compare a multi-population with only GA and PSO populations,
% versus an ensemble of GA and PSO algorithms. We tested on the first five functions of the
% BBOB testbed. We used ten populations and eight workers for the experiment and the
% same parameters as before.

% Shouldn't this be a subsection of experimental setup? - JJ
% \subsection{Hardware}
% \label{sec:hardware}

@@

Using the ECS CLI, we can provision a cluster with a configuration file
specifying the computing resources needed. We can specify the configuration of a Task 
by using a task definition file (see Listing~\ref{code:ecs}) in YAML format.
The main attributes are the {\texttt task\_size} indicating the memory and CPU units
assigned, and network configuration. The complete configuration of the task is generated
by the {\texttt compose command} that uses a {\texttt docker-compose.yml} file to generate 
the configuration on-the-fly. For example to deployed the controller, setup and redis 
containers we used the file in Listing ~\ref{code:compose}.

We deployed two configurations, one that included all containers in a single Task,
and another one, where we assigned each worker to a single Task and the controller, 
redis, and setup in another independent Task.

For the first option, we deployed an AWS Fargate cluster and defined a Task configuration file with
8GBs of memory and 4096 computing units, equivalent to 4 CPUs. Because of the
ten container limit, we only used seven workers and tested function $f_4$ as
in the previous experiment, and to save resources we only tested for dimensions 10 and 20.

For the second option we assigned each worker to a single task with 4GBs of memory and 2048 computing units,
and scale the number of worker tasks to eight.  % Please use caps or not for Tasks in a consistent way - JJ

@@

% Please clarify that those images are published in Docker hub by yourself - JJ
The results are shown in Table X.

>>>>

\usepackage{lineno,hyperref}
\usepackage{rotating}
\usepackage{algpseudocode,algorithm}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{xcolor}
\usepackage[normalem]{ulem}
\useunder{\uline}{\ul}{}

@@

For local experiments we have deployed the EvoSwarm application in a
Desktop PC with an AMD Ryzen 9 3900x 12-core processor with 24 threads with a 3.8 GHz base clock, 
and 16 GB RAM. We used Docker version 19.03.3, build a872fc2f86, and docker-compose version 1.21.0, 
in Ubuntu Linux 18.04; coce is run using Python 3.7.5. Container images and the docker-compose file are available at
\url{https://hub.docker.com/u/mariosky}, and \url{https://github.com/mariosky/EvoSwarm} respectively. 
The BOBB experiments were performed with COCO \cite{hansen2016coco} version bbob.v15.03 in Python, 
the plots were produced with version 2.3.2. % Are we using BBOB plots? - JJ

@@

\begin{figure*}[h!tb]
  \begin{tabular}
      {c@{\hspace*{-0.00001\textwidth}}
       c@{\hspace*{-0.00001\textwidth}}
       c@{\hspace*{-0.00001\textwidth}}
      }
  GA  &  PSO & GA \& PSO\\   
  \includegraphics[width=0.30\textwidth]{GAOnly_f001}&
  \includegraphics[width=0.30\textwidth]{PSOOnly_f001}&
  \includegraphics[width=0.30\textwidth]{GAPSO_f001}\\

  \includegraphics[width=0.30\textwidth]{GAOnly_f002}&
  \includegraphics[width=0.30\textwidth]{PSOOnly_f002}&
  \includegraphics[width=0.30\textwidth]{GAPSO_f002}\\

  \includegraphics[width=0.30\textwidth]{GAOnly_f003}&
  \includegraphics[width=0.30\textwidth]{PSOOnly_f003}&
  \includegraphics[width=0.30\textwidth]{GAPSO_f003}\\

  \includegraphics[width=0.30\textwidth]{GAOnly_f004}&
  \includegraphics[width=0.30\textwidth]{PSOOnly_f004}&
  \includegraphics[width=0.30\textwidth]{GAPSO_f004}\\

  \includegraphics[width=0.30\textwidth]{GAOnly_f005}&
  \includegraphics[width=0.30\textwidth]{PSOOnly_f005}&
  \includegraphics[width=0.30\textwidth]{GAPSO_f005}\\
  \end{tabular}
  \vspace{-3ex}
   \caption{
Scaling of running time with problem dimension to reach certain target values $\Delta f$. Lines:
average runtime (aRT); Cross ($+$): median runtime of successful runs to reach
the most difficult target that was reached at least once (but not always);
Cross (\textcolor{red}{$\times$}): maximum number of f-evaluations in any trial. Notched boxes:
interquartile range with median of simulated runs; all values are divided by
dimension and plotted as $log_{10}$ values versus dimension. 
}
\label{fig:bbob}
\end{figure*}

@@

Having a multi-population based algorithm can decrease the execution time
because of the parallel execution of the evolution. But, having heterogeneous 
populations might enhance evolutionary search and needless
evaluations than homogeneous systems; heterogeneous settings, if
done right, increases the diversity of the whole population\cite{araujo2008multikulti}.

But this is a rule of thumb, and it will depend on the degree of heterogeneity,
as well as on the algorithm itself. Some level of heterogeneity can be
implemented by just changing the configuration parameters of each population,
as in the previous experiment, but in this case, we are interested in a heterogeneous search strategy.

Therefore, in this experiment we compare a multi-population with only GA and PSO populations,
versus an ensemble of GA and PSO algorithms. We tested on the first five functions of the
BBOB testbed. We used ten populations and eight workers for the experiment and the
same parameters as before.

\subsection{Experimental Results}

The results obtained by the experiments show (see Figure \ref{fig:bbob}) how
the runtime scales with dimension to reach certain target values $\Delta f$;
The color of the lines indicates the average runtime, for each target reached
at least one time. A red circle without a number on top indicates the algorithm
has reached the $10^{-8}$ target on all instances for that dimension. If there
is a number on the top, it indicates how many times the algorithm has reached
the target. Fixed values of targets $\Delta f = 10^{k}$ with $k$ colors are
given in the legend of $f_1$, $k$ values are $[-8,-5,-3,-2,-1,0,1]$. Results
from each configuration are presented in each column, GA, PSO, and GA\&PSO
populations.

We can see that the GA algorithm (on the left column) has a hard time reaching even the
$10^{-5}$ target in functions ($f_1-f_4$) for all dimensions, and has the worst
performance in 20 and 40 dimensions. Finally, it performs better on $f_5$
reaching the $10^{-8}$ on all dimensions but with a higher runtime on lower
dimensions.

On the other hand, the multi-population PSO (taking the column on the middle) 
has better performance; its results take a smaller range,
reaching all targets for functions $f_1$ and $f_2$ but with a slightly higher
runtime than the multi-strategy configuration. It is not able to reach the
$10^{-8}$ target on the higher dimensions of $f_3$ and $f_4$. Moreover, the PSO
multi-population has the best runtime for the lower dimensions of $f_5$.

Finally, the proposed PSO\&GA mixed algorithm has reached the $10^{-8}$ target
value on all functions ($f_1-f_5$) and scales well to higher dimensions, even on
$f_3$ and $f_4$ (charts on the lower rows), which are usually considered
difficult functions. This experiment also exemplifies the type of analyses that
we can perform with the EvoSwarm application by reusing the COCO framework.

@@

Using the ECS CLI, we can provision a cluster with a configuration file
specifying the computing resources needed. We can specify the configuration of a task 
by using a task definition file (see Listing~\ref{code:ecs}) in YAML format.
The main attributes are the {\texttt task\_size} indicating the memory and CPU units
assigned, and network configuration. The complete configuration is generated
by the {\texttt ecs compose command} that uses a {\texttt docker-compose.yml} file to generate 
the configuration on-the-fly. For example to deploy the controller, setup and redis 
containers we used the file in Listing ~\ref{code:compose}. We have 
provided the container images and are published in Docker hub \url{https://hub.docker.com/u/mariosky}.

We deployed two configurations, one that included all containers in a single task,
and another one, where we assigned each worker to a different task and the controller, 
redis, and setup in another, for a total of nine concurrent Fargate tasks.

For the first option, we deployed an AWS Fargate cluster and defined a task configuration file with
8GBs of memory and 4096 computing units, equivalent to 4 CPUs. Because of the
ten container limit, we only used seven workers and tested function $f_4$ as
in the previous experiment, and to save resources we only tested for dimensions 10 and 20 for 15 instances.

For the second option, we used the same cluster and additionally
assigned each worker to a task with 4GBs of memory and 2048 computing units. Then scale the 
number of worker tasks to eight. We can scale the number of tasks for up to 100 instances in 
a Fargate configuration. 
                                  % Please use caps or not for Tasks in a consistent way - JJ
                                  % Done -Mario
                                  % Done above - M

@@

\subsection{Experimental Results}

% Please clarify that those images are published in Docker hub by yourself - JJ
The results are shown in Figure~\ref{fig:aws-results} and Table~\ref{tab:aws}.

As we can see, the local environment (see section \ref{setup}) achieved a higher
ratio of evaluations per second (in bold). Still, we must consider that there is a minimum
network overhead in a local workstation and a clock speed of 3.8 GHz. The first
option (labeled Single) is straightforward to deploy, needing only a single
docker-compose file, and one command to execute. But with the limit of 4CPUs and
ten containers, it is the slower option. The second option (labeled Multi), is
faster, and not too far from the local execution, especially when solving the
$f_4$ in 20 dimensions (underline in table). With the capacity for increasing the number of worker
containers, it is reasonable to think that we can achieve a higher number of
evaluations with an appropriate configuration. Nevertheless, the goal of this
experiment was not to increase the number of evaluations but to test the
flexibility of a cloud-native application.

\begin{figure}[h!tbp]
  \centering
  \includegraphics[width=\textwidth]{aws}
  \caption{Boxplot of the number of evaluations per second ratio, for 15 instances of function $f_4$,
   with three execution environments Local, Single task and Multi-task.
   }
  \label{fig:aws-results}
\end{figure}

\begin{table}[]
  \small
  \caption{ Median number of the evaluations per second ratio, for 15 instances of function $f_4$,
  with three execution environments Local, Single task and Multi task}
  \label{tab:aws}
  \vspace{0.25cm}
  \centering

  \begin{tabular}{lll}
  \hline
  Dim & Env         & Median of ratio \\
  \hline
  10  & Local       & \textbf{47,166.29}       \\
      & Multi task  & 40,466.45    \\
      & Single task & 18,193.05      \\
  \hline
  20  & Local       & \textbf{36,101.86}   \\
      & Multi task  & {\ul 33,916.02}       \\
      & Single task & 16,546.95  \\
  \hline
  \end{tabular}
\end{table}
